A thermodynamic view of dusty protoplanetary disks: axisymmetric stability

### 2016-08-28 model.tex
tiny tex and code updates
tiny corrections

We usually consider $\epsilon_0 \ll 1$ and settled dust such that
$\Htilde<\Hgas$. Then $\rhog(r,z)$ is effectively Gaussian, as in the
dust-free case. The dust density is approximately 
\begin{align}
  \rhod \simeq \epsilon_0\rho_\mathrm{g0}(r) \exp
        {\left(-\frac{z^2}{2H_\mathrm{d}^2}\right)}, 
\end{align}
with 
\begin{align}
  \frac{1}{H_\mathrm{d}^2} = \frac{1}{\Htilde^2} + \frac{1}{\Hgas^2}, 
\end{align}
and $H_\mathrm{d}$ is the dust-scale height. In numerical
calculations, we  we specify $H_\mathrm{d}< \Hgas$ to obtain 
$\Htilde$ for input.

@@

\subsection{Vertical buoyancy}\label{vbuoyancy}
Having identified the entropy of an isothermal dusty gas, we find the
vertical buoyancy frequency $N_z$ of the mixture is given by 
\begin{align}
  N_z^2 \equiv - \frac{1}{\rho}\frac{\p P}{\p z}\frac{\p s}{\p z} &=
  \frac{c_s^2(r)}{\left(1+\epsilon\right)^2}\frac{\p\ln\rhog}{\p 
  z}\frac{\p\epsilon}{\p z} \\ &
                                  =
  \frac{\epsilon}{\left(1+\epsilon\right)^2}\left(\frac{z}{\Htilde}\right)^2\OmK^2\notag,  
\end{align}
where the second equality assumes $c_s=c_s(r)$ and the final equality
applies to the equilibria in \S\ref{eqm}. Thus,  
\begin{align*}
N_z\lesssim
O\left(\sqrt{\epsilon_0}\OmK\right). 
\end{align*}
However, for thin dust layers such that $H_\epsilon \ll \Hg$, 
$\mathrm{max}\left(N_z\right)$ may occur outside a finite vertical domain.

>>>>

We usually consider gas-dominated disks with $\epsilon_0 \ll 1$.  
Then $\rhog(r,z)$ is effectively Gaussian, as in the 
dust-free case. The dust density is approximately 
\begin{align}
  \rhod \simeq \epsilon_0\rho_\mathrm{g0}(r) \exp
        {\left(-\frac{z^2}{2H_\mathrm{d}^2}\right)}, 
\end{align}
with 
\begin{align}
  \frac{1}{H_\mathrm{d}^2} = \frac{1}{\Htilde^2} + \frac{1}{\Hgas^2}, 
\end{align}
and $H_\mathrm{d}$ is the dust-scale height. In numerical
calculations, we  we specify $H_\mathrm{d}< \Hgas$ to obtain 
$\Htilde$ for input.

@@

\subsection{Vertical buoyancy}\label{vbuoyancy}
Having identified the entropy of an isothermal dusty gas, we find the
vertical buoyancy frequency $N_z$ of the mixture is given by 
\begin{align}
  N_z^2 \equiv - \frac{1}{\rho}\frac{\p P}{\p z}\frac{\p s}{\p z} &=
  \frac{c_s^2(r)}{\left(1+\epsilon\right)^2}\frac{\p\ln\rhog}{\p 
  z}\frac{\p\epsilon}{\p z} \\ &
                                  =
  \frac{\epsilon}{\left(1+\epsilon\right)^2}\left(\frac{z}{\Htilde}\right)^2\OmK^2\notag,  
\end{align}
where the second equality assumes $c_s=c_s(r)$ and the final equality
applies to the equilibria in \S\ref{eqm}. Thus,  
\begin{align*}
N_z\lesssim
O\left(\sqrt{\epsilon_0}\OmK\right). 
\end{align*}
However, for thick dust layers such that $H_\epsilon \gg \Hg$, 
$\mathrm{max}\left(N_z\right)$ may occur outside a finite vertical domain.

### 2016-08-28 result.tex
minor edits
tiny corrections

For simplicity, we mostly consider constant midplane
dust-to-gas ratios $\epsilon_0$ and its characteristic height
$H_\epsilon$. The latter is specified via the dust layer thickness
$\Hd(r_0)$. (We will allow for radial dependices in $\epsilon_0$ and
$H_\epsilon$ in \S\ref{varHd}.)   
These, together with the perturbation wavenumber $k$, comprises main the
parameters of the linear problem. We fix the radial power-law index
for the midplane gas density to $p = -1.5$, that for the
temperature profile to $q=-1$; and set the gas disk aspect-ratio
$h_\mathrm{g}=0.05$. These are fiducial values used by 
\citet[][\citetalias{lin15} in this section]{lin15}.

@@

Fig. \ref{compare_vshear_varHd} shows that a radially-varying
dust-to-gas ratio increases the magnitude of the vertical shear rate
away from the midplane (see also Eq. \ref{vshear2}). Thus we find
higher VSI growth rates, as shown in Fig. \ref{vsi_dust_varHd}. 
Surfaces mode are more effectively enhanced by the additional vertical 
shear induced by $\p_r\epsilon$. However, all growth rates remain
$O(\hgas\OmK)$. Importantly, the increase in the growth rate of the
fundamental mode is small. Thus we do not expect the radial
dependence in $\epsilon$ to significantly affect the VSI in a dusty
disk.

>>>>

For simplicity, we mostly consider constant midplane
dust-to-gas ratios $\epsilon_0$ and its characteristic height
$H_\epsilon$. The latter is specified via the dust layer thickness
$\Hd(r_0)$. (We will allow for radial dependices in $\epsilon_0$ and
$H_\epsilon$ in \S\ref{varHd}.)   
These, together with the perturbation wavenumber $k$, comprises main the
parameters of the linear problem. We choose a power-law 
midplane gas density profile, $\rho_\mathrm{g0}\propto r^q$ with $p =
-1.5$. The power-law index for the 
temperature profile is $q=-1$; and we set the gas disk aspect-ratio
$h_\mathrm{g}=0.05$. These are fiducial values used by 
\citet[][\citetalias{lin15} in this section]{lin15}.

@@

Fig. \ref{compare_vshear_varHd} shows that a radially-varying
dust-to-gas ratio increases the magnitude of the vertical shear rate
away from the midplane (see also Eq. \ref{vshear2}). Thus we find
higher VSI growth rates, as shown in Fig. \ref{vsi_dust_varHd}. 
Surfaces modes are more effectively enhanced by the additional vertical 
shear induced by $\p_r\epsilon$. However, all growth rates remain
$O(\hgas\OmK)$. Importantly, the increase in the growth rate of the
fundamental mode is small. Thus we do not expect the radial
dependence in $\epsilon$ to significantly affect the VSI in a dusty
disk.
